Different rules for binocular combination of luminance flicker in cortical and subcortical pathways

### 2023-06-22 manuscript.Rmd
Trying again

Classic studies investigating the neurophysiological architecture of V1 reported that cells in cytochrome-oxidase 'blobs' [@Horton1981; @Livingstone1984] are biased towards low spatial frequencies [@Tootell1988; @Edwards1995], and relatively insensitive to stimulus orientation [@Horton1981; @Livingstone1984; though see @Economides2011]. As the blob regions are embedded within ocular dominance columns [@Horton1981], they are also largely monocular [@Livingstone1984; @Tychsen2004]. More recent work has reported psychophysical evidence for unoriented chromatic [@Gheiratmand2013] and achromatic [@Meese2011] mechanisms, that also appear to be monocular. Our use of luminance flicker might preferentially stimulate these mechanisms, perhaps explaining why our EEG data show little evidence of binocular interactions. Indeed, our EEG results could potentially be explained by a model involving entirely non-interacting monocular channels, with the binocular facilitation effects we find (e.g. Figures \@ref(fig:BSratios) & \@ref(fig:TFdata)) owing to additivity of the electrophysiological response across independent monocular cells, rather than within binocular neurons. In the steady-state literature, one hallmark of a nonlinear system is the presence of intermodulation responses at the sums and differences of fundamental flicker frequencies [@Tsai2012]. In Figure \@ref(fig:IMfig) we plot the amplitude spectra of conditions from Experiment 1 in which the two eyes were stimulated at different frequencies (2Hz and 1.6Hz) but at the same contrast (48\%; these correspond to the binocular cross and dichoptic cross conditions in Figures \@ref(fig:pupildata)d,e and \@ref(fig:EEGdata)d,e). Consistent with the temporal properties of pupil responses and EEG, Figure \@ref(fig:IMfig)a reveals a strong intermodulation difference response at 0.4Hz (red dashed line), and Figure \@ref(fig:IMfig)b reveals an intermodulation sum response at 3.6Hz (red dashed line). The presence of these intermodulation terms is predicted by nonlinear gain control models of the type considered here [@Tsai2012; @Baker2017], and indicates that the processing of monocular flicker signals is not fully linear prior to the point at which they are combined across the eyes.

>>>>

Classic studies investigating the neurophysiological architecture of V1 reported that cells in cytochrome-oxidase 'blobs' [@Horton1981; @Livingstone1984] are biased towards low spatial frequencies [@Tootell1988; @Edwards1995], and relatively insensitive to stimulus orientation [@Horton1981; @Livingstone1984; though see @Economides2011]. As the blob regions are embedded within ocular dominance columns [@Horton1981], they are also largely monocular [@Livingstone1984; @Tychsen2004]. More recent work has reported psychophysical evidence for unoriented chromatic [@Gheiratmand2013] and achromatic [@Meese2011] mechanisms, that also appear to be monocular. Our use of luminance flicker might preferentially stimulate these mechanisms, perhaps explaining why our EEG data show little evidence of binocular interactions. Indeed, our EEG results could potentially be explained by a model involving entirely non-interacting monocular channels, with the binocular facilitation effects we find (e.g. Figures \@ref(fig:BSratios) \& \@ref(fig:TFdata)) owing to additivity of the electrophysiological response across independent monocular cells, rather than within binocular neurons. We therefore performed an additional analysis to investigate this possibility.

In the steady-state literature, one hallmark of a nonlinear system that pools inputs is the presence of intermodulation responses at the sums and differences of the fundamental flicker frequencies [@Tsai2012]. In Figure \@ref(fig:IMfig) we plot the amplitude spectra of conditions from Experiment 1 in which the two eyes were stimulated at different frequencies (2Hz and 1.6Hz) but at the same contrast (48\%; these correspond to the binocular cross and dichoptic cross conditions in Figures \@ref(fig:pupildata)d,e and \@ref(fig:EEGdata)d,e). Consistent with the temporal properties of pupil responses and EEG activity, Figure \@ref(fig:IMfig)a reveals a strong intermodulation difference response at 0.4Hz (red dashed line), and Figure \@ref(fig:IMfig)b reveals an intermodulation sum response at 3.6Hz (red dashed line). The presence of these intermodulation terms is predicted by nonlinear gain control models of the type considered here [@Tsai2012; @Baker2017], and indicates that the processing of monocular flicker signals is not fully linear prior to the point at which they are combined across the eyes. Indeed, our model architecture makes specific predictions about the location of interocular suppression - it impacts before binocular combination, consistent with results from primate physiology [@].
